\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-05-02 trunk/related.tex
updates

Such techniques have proved their mettle in the first round of use of
AUVs; they've provided operators a simple way to handle control
complexity of the vehicle while returned data at far higher resolution
than would have been possible using traditional ship-based methods and
at substantially reduced amortized costs of platform operation and
ownership over multiple years. AUV operators and clients for their
data products have reasons to be well satisfied with results thus
far. However, as hardware and more cost-effective sensors become more
robust to the environment and as mission durations increase (as
demonstrated by glider experiments such as \cite{rucool11}), sustained
presence in the ocean requires the ability to deal with off-nominal
conditions, that balances the needs of high-level mission management,
low-level navigation, instrument control and detection of unstructured
and poorly understood phenomena. Earlier reactive methods are unable
to perform such a balancing act without skewering the foundations of
the software development methodology leading to code bloat or worse.
\cite{carreras06} provides a reasonably good overview of AUV control
architectures that essentially is based on reactive control methods.

A word about generalized planning. Traditionally planning has been
considered computationally intensive (the general planning problem can
be worse than NP-complete \cite{ghallab04}). A major reason for this
belief had to do with the role of the planner (which was assumed to be
generative) and its place in the architecture (infrequently called to
re-plan the \emph{entire} mission in off-nominal conditions) within a
\emph{sense-plan-act} paradigm. This can limit the reactivity of a
robot especially when the environment could change at a rate faster
than the planner can plan.

The architecture of a typical AI planner is along the lines of
Fig. \ref{fig:planner}. Given a \emph{domain model} which encodes the
platform constraints in a higher-level language, an \emph{initial
  state}, a high-level objective or end \emph{goal}, the \emph{plan
  database} (or \texttt{plandb}) is where all assertions are tracked
and maintained. Theorem proving in the form of satisfaction of axioms
embodied in the domain model occurs within the \texttt{plandb}. It is
the job of the \emph{search engine} to provide the inferential
mechanism for state space exploration potentially aided by heuristics
encoded for \emph{search control}. In some complex domains
\cite{mus98} \emph{external modules} can provide additional domain
expertise which can augment the domain model. The end result is plan
formulation using a multitude of approaches either using forward or
backward search or using more generalized \emph{partial order} methods
\cite{ghallab04}.

@@

@@

Planning and plan execution are not new to the overall field of
robotics. Early motivation of building computational mechanisms for
decision making were intended to be deployed on robotic devices by
very definition. The seminal volume \cite{computersthought} very
clearly articulates how physical manifestation of robots could be
controlled by task planning and execution. The very first planner
\cite{green69} was an exercise in computational theorem proving with
early state space planners popularized by \texttt{STRIPS}
\cite{strips71} dominating the academic landscape. An early
demonstration of planning as a situated agent\footnote{\kcomment{By
    \emph{situated} we emphasize that an agent is embedded within a
    physical robot.}} was the highly influential \texttt{Shakey} the
robot experiment \cite{shakey84}. Subsequent work driven largely by
defense funding in the United States sequestered the discipline in
academic labs principally applied for algorithmic development and if
embedded then in testing within the confines of modest indoor
settings. The leap to applying them for real world problem solving is
relatively new and driven largely by NASA applications pioneered by
\cite{mus94,mus98, jonsson00, rajan00, chien05, bresina05}.

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale), notable examples of which include \texttt{IPEM}
\cite{AmbrosIngerson88}, \texttt{ROGUE} \cite{Haigh98}, the LAAS
Architecture \cite{alami:1998p820}, the Remote Agent Experiment
\texttt{RAX} \cite{mus98} and the Autonomous Spacecraft Experiment
\texttt{ASE} \cite{chien99} (see \cite{Knight01} for a survey). % Early
% efforts were cognizant of if absent with, dealing with time and
% resources for such experiments.

Representationally, the first paper on managing time systematically
within AI Planning frameworks was \cite{dean87}; subsequent efforts by
Boddy and others \cite{Dean88,Boddy93} refined the notion of using a
temporal database and the conceptualization of temporal
intervals. Using these base-level concepts Muscettola and others
\cite{mus92, ghallab94, laborie95, cesta96} developed the notion of
plan/schedule envelopes using the notion of state variable
instantiated timelines. Simultaneously work by Dechter et.al
\cite{dechter91} resulted in efficient temporal constraint propagation
which systematically defined the notion of Simple Temporal Networks
(STNs) and path and arc consistency algorithms. This work neatly put
together earlier efforts by Mackworth and Freuder \cite{mackworth77,
  mack85} and the Constraints community on propagation algorithms. In
parallel work in the UK as a result of US DARPA funding, resulted in
the design of \texttt{O-Plan} \cite{currie91}, which leveraged the
notion of constraints and metric time, but within a Hierarchical Task
Network (HTN) representation. All these efforts were soon after Vere’s
paper on planning and time \cite{vere83} which had a profound effect
on the AI Planning community. In particular the work of Muscettola
\cite{mus94} was embraced by NASA for telescope scheduling. Around the
same time period, researchers at LAAS came up with an architecture
\cite{alami:1998p820} which encapsulated the \texttt{IxTeT} planner
\cite{ghallab94}. It featured advanced concepts for a temporal planner
with the notion of chronicles, plan recognition for partial plan
fragments and early use of systematic resource
constraints. Development on \texttt{IxTeT} has continued albeit at a
slower pace with the contribution of various PhD students’ thesis.

@@

While \texttt{HSTS} \cite{mus94} was being implemented as a
ground-based planning tool for decision support for US based space
observations (EUVE and Cassini) \cite{mus95}, the opportunity to fly
onboard NASA’s New Millennium Deep Space $1$ came about. The design of
the Remote Agent Experiment \texttt{RAX} \cite{pell97, bernard98,
  pell98, mus98, DS1report, rajan00, jonsson00} was a direct off-shoot
of this effort where \texttt{HSTS} (written in LISP) was flown on
board and successfully commanded the spacecraft for a week in May
1999. There were a number of significant software engineering lessons
learned with the HSTS technology as integrated for the RAX:

@@

\item Flexible temporal representation generating a range of plans as
  against a single plan, were robust for plan execution.

@@

It was with the demonstration of \texttt{RAX} $65$ Million miles in
outer space, that temporal reasoning using \kcomment{planning} methods
came to the forefront of AI Planning for situated agents. Lessons
learned from \texttt{RAX} led to the development of \texttt{IDEA}
\cite{mus02,mus04,Dias:2003ua,mus06} with the central theme of using
planners collectively for problem solving arose. Another critical
theme was iteratively and incremental plan repair of an anytime plan
\cite{Zaimag96} as proposed by \texttt{CASPER} \cite{chien00}.

This last lesson in particular had a lasting impact with the
observation that planning and execution are strongly
inter-related. This core concept was behind the next generation of the
Remote Agent, called \texttt{IDEA} (Intelligent Deployable Execution
Agents) where planning and execution were \emph{intertwined} within a
single domain model that spanned the most abstract (for high-level
planning) to the least (for diagnosis). \texttt{IDEA} \cite{mus02,
  mus04} agents were expected to interact to achieve plan formulation;
however there was no systematic framework for formally governing these
interactions. \texttt{IDEA} was also computationally heavy and
required substantial effort to customize for an application domain. It
did move to a retooled version of \texttt{HSTS}, now called EUROPA
\cite{frank2003, barreiro09} and ported to the C++ language.

While \texttt{IDEA} was being deployed as a coordinating executive for
earth analog rover deployments \cite{wetter05} a separate development
was undertaken to deploy the EUROPA planner for NASA’s Mars
Exploration Rovers (MER) mission. The \texttt{MAPGEN}
(Mixed-initiative Activity Plan GENerator) planner \cite{bresina03,
  aichang04, bresina05, bresina05a} as a decision support tool in the
mixed-initiative mode; \texttt{MAPGEN} continues to be used to this
day to command the twin rovers on the Red
Planet. % While EUROPA2 was not
% designed as an embedded planner, NASA undertook a bottom-up
% re-implementation of the planner and substantially increased its
% performance. EUROPA2 was then released as open source to the AI
% community at large [Europa].

The Autonomous Sciencecraft Experiment (\texttt{ASE}) \cite{chien99,
  chien03} was another autonomy demonstration in space, this time for
a more observable vehicle, the EO-1 in Earth orbit. It encapsulated a
classical 3-layered architecture like \texttt{RAX}. The fundamental
functional difference was that goals could not only be sent by a
ground segment, but also by an onboard science driver which could
encapsulate a range of Machine Learned feature detectors which could
trigger the planner. However ASE was not an integrated system; the
\texttt{CASPER} planner \cite{chien00} would generate plans which were
executed by a separate commercially available rule-based system
\cite{icl}. \texttt{CASPER} touts a \emph{continuous} paradigm for
planning; however this incremental modification simply takes the
overall mission plan and selectively plans near-term at a more
detailed level. The system also uses iterative-repair as a way to fix
plans; however this too is derived from a model of the mission and
spacecraft constraints which allow simple reconfiguration modes
towards replanning. Robust software engineering is also not an
important aspect targeted by \texttt{ASE} given the disparate models
between the planner and the rule-based executive. With such a system,
real-time updates from the environment cannot be accommodated since
the planner is a monolithic computational environment. The rate of
updates from the science drivers is roughly \kcomment{on} the order of
an earth orbit on the EO-1 vehicle. As a result, the ASE is only an
incremental update on RAX and is not suitable for domains where
complex automated reasoning and asynchronous environmental
observations need to be factored in deliberation.

\texttt{TCA} \cite{simmons94} is another framework for robot control
and has similar motivations to \texttt{RAX} and
\texttt{ASE}. Reactivity and deliberation are also key to \texttt{TCA}
and a systematic approach to development led to its design. However it
suffers from three key weaknesses. First it has a weaker
representational framework of Task Trees which require representation
of tasks within cleanly formulated hierarchies. Constraints between
branches within a task hierarchy are allowed; timing between leaves of
different trees is however not possible. Second, its temporal
framework does not deal with flexibility; time points are fixed and
while partial orders are possible, it does deal with execution time
uncertainty. Third, while \texttt{TCA} shares the notion of using
domain expertise of different planners, it does so thru a centralized
message passing mechanism. While this allows disparate code-bases to
talk to each other \kcomment{through} IPC (\kcomment{an} inter-process
communication or message passing mechanism), failure in the
centralized controller is tantamount to a system crash.

The \texttt{ReSSAC} project at ONERA \cite{teichteil07} has similar
motivations in controlling aerial UAV platforms for autonomous Search
and Rescue. Like \texttt{TCA}, the project uses a supervisor to
control planning and at the same time decouples the deliberation and
reactive components from the lower level functional layer. This
deliberate separation is partly because of the use of optimizing MDP’s
(Markov Decision Processes \cite{mdp93}) for planning which uses
probabilistic state transitions to determine the next course of action
based on perceived state. To overcome the typical problem of state
space explosion with MDPs, a local heuristic is used in order to
generate reachable goal states incrementally. This work however does
not deal with metric time with a modestly simple model.

\texttt{CIRCA} \cite{musliner95} is another effort to bring decision
making for real-world problems with the augmented need to have
\emph{verifiable} controllers synthesized in-situ to ensure that
within explored states, there are no erroneous transitions to
dangerous states by asynchronously generating Test-Action-Pairs
(TAPs).  These are annotated production rules consisting of a set of
tests (or preconditions) and a set of actions to take if all the tests
return true. TAPs are synthesized and scheduled by \texttt{CIRCA} and
provide a viable way to deal with a dynamically changing
world. \texttt{CIRCA} however differentiates between reasoning about
time and reasoning in real-time with the implication that reasoning
necessarily requires substantial computation for state space
exploration which negates the real-time (and thus real-world)
impact.

A key and early concern that dominated planning was that of
scalability.  The planning cycle in many approaches was monolithic
often making fast reaction times impractical when necessary for
embedded robotic agents. Many of these systems (which were
$3$-layered) also utilized very different techniques for specifying
each layer in the architecture resulting in duplication of effort and
a diffusion of knowledge.  The work we highlight in this chapter
builds on the approach used by \texttt{IDEA} \cite{mus02, mus04} in
utilizing a collection of controllers, each interleaving planning and
execution within a common framework. \texttt{IDEA} however, provided
no support for conflict resolution between controllers, nor does it
provide an efficient algorithm for integrating current state within a
controller, relying instead on a possibly exponential planning
algorithm. Efficient synchronization of state in a partitioned
structure is fundamental to making the approach effective in practice.

Our system, \rxe, the focus of this chapter, is an agent control
structure formally defined as a composition of coordinated control
loops where sensing, planning and acting are resulting from concurrent
deliberating tasks within a formal framework.  \rx was designed after
carefully evaluating lessons learned from \texttt{IDEA} to which it
owes substantially the notion of partitioned problem solving. However
partitioning in \rx is systematic and methodologically principled
where we manage the information flow within the partitioned structure
to ensure consistency in order to direct the flow of goals and
observations in a timely manner. The resulting control structure
improves scalability since many details of each controller can be
encapsulated within a single control loop.  Furthermore, partitioning
increases robustness since controller failure can be localized to
enable graceful system degradation, making this an effective
\emph{divide-and-conquer} approach to the overall control problem.

A recent variation of the key idea of such controller composition
derived from \rx is \texttt{ROAR} \cite{degroote11} where lower-level
functional modules that control specific sensors are organized within
a graph hierarchy. One reason they cite such a structure is essential
is in dealing robustly to failure; an off-nominal condition will allow
rapid graph traversal to identify alternative ways in which sensing
tasks can be executed and therefore aiding in responsiveness. However,
it is unclear whether to date this system actually deliberates and
whether it is actually instantiated on a real platform for control.

>>>>

Such techniques have proved their mettle in the first round of use of
AUVs; they've provided operators a simple way to handle control
complexity of the vehicle while returned data at far higher resolution
than would have been possible using traditional ship-based methods and
at substantially reduced amortized costs of platform operation and
ownership over multiple years. AUV operators and clients for their
data products have reasons to be well satisfied with results thus
far. However, as hardware and more cost-effective sensors become more
robust to the environment and as mission durations increase (as
demonstrated by glider experiments such as \cite{rucool11}), sustained
presence in the ocean requires the ability to deal with off-nominal
conditions, that balances the needs of high-level mission management,
low-level navigation, instrument control and detection of unstructured
and poorly understood phenomena. Earlier reactive methods are unable
to perform such a balancing act without skewering the foundations of
the software development methodology leading to code bloat or worse.
\cite{carreras06} provides a good overview of AUV control
architectures essentially based on reactive control methods.

@@

Traditionally generalized planning has been considered computationally
intensive (the general planning problem can be worse than NP-complete
\cite{ghallab04}). A major reason for this belief had to do with the
role of the planner (which was assumed to be generative) and its place
in the architecture (infrequently called to re-plan the \emph{entire}
mission in off-nominal conditions) within a \emph{sense-plan-act}
paradigm. This can limit the reactivity of a robot especially when the
environment could change at a rate faster than the planner can plan.

The architecture of a typical AI planner is along the lines of
Fig. \ref{fig:planner}. Given a \emph{domain model} which encodes the
platform constraints in a higher-level language, an \emph{initial
  state}, a high-level objective or end \emph{goal}, the \emph{plan
  database} (or \texttt{plandb}) is where all assertions are tracked
and maintained. Theorem proving in the form of satisfaction of axioms
embodied in the domain model occurs within the \texttt{plandb}. It is
the job of the \emph{search engine} to provide the inferential
mechanism for state space exploration potentially aided by heuristics
encoded for \emph{search control}. In some complex domains
\cite{mus98} \emph{external modules} can provide additional domain
expertise which can augment the model. The end result is plan
formulation using a multitude of approaches either using forward or
backward search or using more generalized \emph{partial order} methods
\cite{ghallab04}.

@@

Planning and plan execution are not new to the overall field of
robotics. Early motivation of building computational mechanisms for
decision making were intended to be deployed on robotic devices by
very definition. The seminal volume \cite{computersthought} very
clearly articulates how physical manifestation of robots could be
controlled by task planning and execution. The very first planner
\cite{green69} was an exercise in computational theorem proving with
early state space planners popularized by \texttt{STRIPS}
\cite{strips71} dominating the academic landscape. An early
demonstration of planning as a situated agent\footnote{\kcomment{By
    \emph{situated} we emphasize that an agent is embedded within a
    physical robot.}} was the highly influential \texttt{Shakey} the
robot \cite{shakey84}. Subsequent work driven largely by defense
funding in the United States sequestered the discipline in academic
labs principally applied for algorithmic development and if embedded
then in testing within the confines of \kcomment{structured} indoor
settings. The leap to applying them for real world problem solving is
relatively new and driven largely by NASA applications pioneered by
\cite{mus94,mus98, jonsson00, rajan00, chien05, bresina05}.

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale \kcomment{and survey}), notable examples of which include
\texttt{IPEM} \cite{AmbrosIngerson88}, \texttt{ROGUE} \cite{Haigh98},
the LAAS Architecture \cite{alami:1998p820}, the Remote Agent
Experiment \texttt{RAX} \cite{mus98} and the Autonomous Spacecraft
Experiment \texttt{ASE} \cite{chien99} (see \cite{Knight01} for a
survey).

Representationally, the first paper on managing time systematically
within AI Planning frameworks was \cite{dean87}; subsequent efforts by
Boddy and others \cite{Dean88,Boddy93} refined the notion of using a
temporal database and the conceptualization of temporal
intervals. Using these concepts Muscettola and others \cite{mus92,
  ghallab94, laborie95, cesta96} developed the notion of plan/schedule
envelopes using the notion of state variable instantiated
timelines. Simultaneously work by Dechter et.al \cite{dechter91}
resulted in efficient temporal constraint propagation which
systematically defined the notion of Simple Temporal Networks (STNs)
and path and arc consistency algorithms. This work neatly put together
earlier efforts by Mackworth and Freuder \cite{mackworth77, mack85}
and the Constraints community on propagation algorithms. In parallel
work in the UK, resulted in the design of \texttt{O-Plan}
\cite{currie91}, which leveraged the notion of constraints and metric
time, but within a Hierarchical Task Network (HTN) representation.
These efforts were soon after Vere’s paper on planning and time
\cite{vere83} which had a profound effect on the AI Planning
community. In particular the work of Muscettola \cite{mus94} was
embraced by NASA for telescope scheduling. Around the same time
period, researchers at LAAS came up with an architecture
\cite{alami:1998p820} which encapsulated the \texttt{IxTeT} planner
\cite{ghallab94} \kcomment{which dealt with time and resources}. It
featured advanced concepts for a temporal planner with the notion of
chronicles, plan recognition for partial plan fragments and early use
of systematic resource constraints. % Development on \texttt{IxTeT} has
% continued albeit at a slower pace with the contribution of various PhD
% students’ thesis.

@@

While \texttt{HSTS} \cite{mus94} was being implemented as a
ground-based planning tool for decision support for US based space
observations (EUVE and Cassini) \cite{mus95}, the opportunity to fly
onboard NASA’s New Millennium Deep Space $1$ came about. The design of
the Remote Agent Experiment \texttt{RAX} \cite{pell97, bernard98,
  pell98, mus98, DS1report, rajan00, jonsson00} was a direct off-shoot
of this effort where \texttt{HSTS} (written in LISP) was flown on
board and successfully commanded the spacecraft for a week in May
1999\footnote{This was the first (and to our knowledge only) software
  ever to been written in \texttt{LISP} to be flown in space.}. There
were a number of significant software engineering lessons learned with
the HSTS technology as \kcomment{implemented} for the RAX
\kcomment{experiment}:

@@

\item Flexible temporal representation generating a range of plans as
  against a single plan, were robust for \kcomment{robotic} plan
  execution.

@@

It was with the demonstration of \texttt{RAX} $65$ Million miles in
outer space, that temporal reasoning using \kcomment{planning} methods
came to the forefront of AI Planning for situated agents. Lessons
learned from \texttt{RAX} led to the development of \texttt{IDEA}
\cite{mus02,mus04,Dias:2003ua,mus06} with the central theme of using
planners collectively for problem solving. Another critical theme was
iteratively and incremental plan repair of an anytime plan
\cite{Zaimag96} as proposed by \texttt{CASPER} \cite{chien00}.

This last lesson in particular had a lasting impact with the
observation that planning and execution are strongly
inter-related. This core concept was behind the next generation of the
Remote Agent, called \texttt{IDEA} (Intelligent Deployable Execution
Agents) where planning and execution were \emph{intertwined} within a
single domain model that spanned the most abstract (for high-level
planning) to the least (for diagnosis). \texttt{IDEA} \cite{mus02,
  mus04} agents were expected to interact to achieve plan formulation;
however there was no systematic framework for formally governing these
interactions. \texttt{IDEA} was also computationally heavy and
required substantial effort to customize for an application domain. It
did move to a retooled version of \texttt{HSTS}, called \eu
\cite{frank2003, barreiro09} while being ported to the C++ language.

While \texttt{IDEA} was being deployed as a coordinating executive for
earth analog rover deployments \cite{wetter05} a separate development
was undertaken to deploy the \eu planner for NASA’s Mars Exploration
Rovers (MER) mission. The \texttt{MAPGEN} (Mixed-initiative Activity
Plan GENerator) planner \cite{bresina03, aichang04, bresina05,
  bresina05a} as a decision support tool in the mixed-initiative mode;
\texttt{MAPGEN} continues to be used to this day to command the twin
rovers on the Red Planet. % While EUROPA2 was not
% designed as an embedded planner, NASA undertook a bottom-up
% re-implementation of the planner and substantially increased its
% performance. EUROPA2 was then released as open source to the AI
% community at large [Europa].

The Autonomous Sciencecraft Experiment (\texttt{ASE}) \cite{chien99,
  chien03} was another autonomy demonstration in space, this time for
a more observable vehicle, the EO-1 in Earth orbit. It encapsulated a
classical 3-layered architecture like \texttt{RAX}. Goals could not
only be sent by a ground segment, but also by an onboard science
driver which could encapsulate a range of Machine Learned feature
detectors which could trigger the planner. However \texttt{ASE} was
not an integrated system; the \texttt{CASPER} planner \cite{chien00}
would generate plans which were executed by a separate commercially
available rule-based system \cite{icl}. \texttt{CASPER} demonstrates
what is called as a \emph{continuous} paradigm for incremental
planning and iterative plan repair; this methodology takes the overall
mission plan and selectively plans \kcomment{for the} near-term at a
more detailed level.
% The system also uses iterative-repair as a way to fix
% plans; however this too is derived from a model of the mission and
% spacecraft constraints which allow simple reconfiguration modes
% towards replanning. 
Robust software engineering is also not an important aspect targeted
by \texttt{ASE} given the disparate models between the planner and the
rule-based executive. With such a system, real-time updates from the
environment could not be accommodated since the planner is a
monolithic computational environment. The rate of updates from the
science drivers is roughly \kcomment{on} the order of an earth orbit
on the EO-1 vehicle. As a result, the \texttt{ASE} is only an
incremental update on \texttt{RAX} and is not suitable for domains
where complex automated reasoning and asynchronous environmental
observations need to be factored in deliberation.

\texttt{TCA} \cite{simmons94} another framework for robot control and
has similar motivations to \texttt{RAX} and \texttt{ASE}. Reactivity
and deliberation are also key to \texttt{TCA} \kcomment{along with} a
systematic approach to development led to its design. However it
suffers from three key weaknesses. First it has a weaker
representational framework of Task Trees which require representation
of tasks within cleanly formulated hierarchies. Constraints between
branches within a task hierarchy are allowed; timing between leaves of
different trees is however not possible. Second, its temporal
framework does not deal with flexibility; time points are fixed and
while partial orders are possible, it does deal with execution time
uncertainty. Third, while \texttt{TCA} shares the notion of using
domain expertise of different planners, it does so thru a centralized
message passing mechanism. While this allows disparate code-bases to
talk to each other \kcomment{through} IPC (\kcomment{an} inter-process
communication or message passing mechanism), failure in the
centralized controller is tantamount to a system crash.

The \texttt{ReSSAC} project at ONERA \cite{teichteil07} \kcomment{is
  motivated by} controlling aerial UAV platforms for autonomous Search
and Rescue. Like \texttt{TCA}, the project uses a supervisor to
control planning and at the same time decouples the deliberation and
reactive components from the lower level functional layer. This
deliberate separation is partly because of the use of optimizing MDP’s
(Markov Decision Processes \cite{mdp93}) for planning which uses
probabilistic state transitions to determine the next course of action
based on perceived state. To overcome the typical problem of state
space explosion with MDPs, a local heuristic is used in order to
generate reachable goal states incrementally. This work however does
not deal with metric time \kcomment{and comes} with a modestly simple
model.

\texttt{CIRCA} \cite{musliner95} is an effort to bring decision making
for real-world problems with the augmented need to have
\emph{verifiable} controllers synthesized in-situ to ensure that
\kcomment{during state space exploration}, there are no erroneous
transitions to dangerous states by asynchronously generating
Test-Action-Pairs (TAPs).  These are annotated production rules
consisting of a set of tests (or preconditions) and a set of actions
to take if all the tests return true. TAPs are synthesized and
scheduled by \texttt{CIRCA} and provide a viable way to deal with a
dynamically changing world. \texttt{CIRCA} however differentiates
between reasoning about time and reasoning in real-time with the
implication that reasoning necessarily requires substantial
computation for state space exploration which negates the real-time
(and thus real-world) impact.

A key and early concern that dominated planning was that of
scalability.  The planning cycle in many approaches was \kcomment{(and
  is)} monolithic, often making fast reaction times impractical when
necessary for embedded robotic agents. Many of these systems (which
were $3$-layered) also utilized very different techniques for
specifying each layer in the architecture resulting in duplication of
effort and a diffusion of knowledge.  The work we highlight in this
chapter builds on the approach used by \texttt{IDEA} \cite{mus02,
  mus04} in utilizing a collection of controllers, each interleaving
planning and execution within a common framework. \texttt{IDEA}
however, provided no support for conflict resolution between
controllers, nor did it provide an efficient algorithm for integrating
current state within a controller, relying instead on a possibly
exponential planning algorithm. Efficient synchronization of state in
a partitioned structure is fundamental to making the approach
effective in practice.

Our system, \rxe, the focus of this chapter, is an agent control
structure formally defined as a composition of coordinated control
loops where sensing, planning and acting \kcomment{result} from
concurrent deliberating tasks within a formal framework.  \rx was
designed after carefully evaluating lessons learned from \texttt{IDEA}
to which it owes substantially the notion of partitioned problem
solving. Partitioning in \rx is systematic and methodologically
principled where we manage the information flow within the partitioned
structure to ensure consistency in order to direct the flow of goals
and observations in a timely manner. The resulting control structure
improves scalability since many details of each controller can be
encapsulated within a single control loop.  Furthermore, partitioning
increases robustness since controller failure can be localized to
enable graceful system degradation, making this an effective
\emph{divide-and-conquer} approach to \kcomment{solving} the overall
control problem.

A recent variation of the key idea of such controller composition
derived from \rx is \texttt{ROAR} \cite{degroote11} where lower-level
functional modules that control sensors are organized within a graph
hierarchy. One reason they cite such a structure \kcomment{as being}
essential is in dealing robustly to failure; an off-nominal condition
will allow rapid graph traversal to identify alternative ways in which
sensing tasks can be executed and therefore aiding \kcomment{platform}
responsiveness. % However,
% it is unclear whether to date this system actually deliberates and
% whether it is actually instantiated on a real platform for control.

### 2012-05-02 trunk/setobook.tex
updates

\abstract{We describe a general purpose Artificial Intelligence based
  control architecture that incorporates in-situ decision making for
  autonomous underwater vehicles (AUVs). The Teleo-Reactive EXecutive
  (\rx) framework deliberates about future states, plans for actions
  and executes generated activities while monitoring plans for
  anomalous conditions. Plans are no longer scripted a priori but
  synthesized onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation. \rx is in routine operational use at MBARI, providing
  scientists a new tool to sample and observe the dynamic coastal
  ocean.}

>>>>

\abstract{We describe a general purpose Artificial Intelligence based
  control architecture that incorporates in-situ decision making for
  autonomous underwater vehicles (AUVs). The Teleo-Reactive EXecutive
  (\rxe) framework deliberates about future states, plans for actions
  and executes generated activities while monitoring plans for
  anomalous conditions. Plans are no longer scripted a priori but
  synthesized onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation. \rx is in \kcomment{regular} operational use at MBARI,
  providing scientists a new tool to sample and observe the dynamic
  coastal ocean.}
